---
title: "Dplyr tutorial"
format: html
editor: visual
execute:
  echo: false
  warning: false
  keep-md: true
---

```{r}
library(here)
library(dplyr)
library(tidyr)
library(knitr)
library(ggplot2)
here::i_am("gitdplyr.Rproj")
raw_spotify <- readRDS('top-spotify-hits-2000-2019.Rds')
```

## Exercise 1

### Question 1

```{r}
spotify <- raw_spotify |> distinct(song, artist, .keep_all = TRUE)
```

The dataset contains:\
- `r n_distinct(spotify$song, spotify$artist)` different songs (`r n_distinct(spotify$song)` different titles)\
- `r n_distinct(spotify$artist)` different artists\
- `r n_distinct(spotify$genre)` different genres.

### Question 2

```{r}
artists_per_year <- spotify |>
  group_by(year) |>
  summarise('Nb artists' = n_distinct(artist))

kable(artists_per_year, format = "markdown")
```

### Question 3

```{r}
most_popular_artist <- spotify |>
  group_by(artist) |>
  summarise(NumSongs = n()) |>
  slice_max(order_by = NumSongs)
```

The most popular artist is `r most_popular_artist$artist` with `r most_popular_artist$NumSongs` songs.

### Question 4

```{r}
tempo_sum <- spotify |>
  group_by(genre) |>
  summarise('min'=min(tempo), 
            'max'=max(tempo), 
            'mean'=mean(tempo),
            'median'=median(tempo),
            'number'=n())

kable(tempo_sum, format = "markdown", digits = 1, align = "lccccc")
```

### Question 5

```{r}
live_n_dance <- spotify |>
  group_by(year) |>
  summarise(Mean_live = mean(liveness), Mean_dance = mean(danceability))

kable(live_n_dance, format = "markdown", digits = 3)
```

### Question 6

```{r}
#| label: fig1-means
graph <- live_n_dance |>
  gather(key = var1, value = val1, Mean_live, Mean_dance)

ggplot(graph, aes(x=year, y = val1))+
  geom_line(aes(color=var1)) +
  scale_color_manual(values = c("darkred", "darkblue"), name = 'Variables', labels = c('Danceability', 'Liveness')) +
  labs(x= 'Year', y= 'Mean') +
  theme(legend.position="bottom")
```

## Exercice 2

```{r}
dropout <- readRDS('dropout.Rds')
```

### Question 1

```{r}
adminGrade <- dropout |>
  group_by(Target, `Marital status`) |>
  summarise(medianAG = median(`Admission grade`))

kable(adminGrade, format = "markdown")
```

### Question 2

```{r}
adminGradeTable <- adminGrade |>
  pivot_wider(names_from = Target, values_from = medianAG)

kable(adminGradeTable, format = "markdown", digits = 2, align = "lccc")
```

### Question 3

```{r}
medianCU <- dropout |>
  group_by(Gender) |>
  summarise(across(starts_with("Curricular"), median))

kable(medianCU, format = "markdown")
```

### Question 4

```{r}
medianCUTable <- medianCU |>
  pivot_longer(cols = starts_with("Curricular")) |>
  pivot_wider(names_from = Gender)

kable(medianCUTable, format = "markdown", digits = 1, align = "lcc", col.names = c('Units', 'Male', 'Female'))
```
